Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks

### 2022-12-29 paper-1/manuscript.tex
Rev Fig6 legend & text (added panel e)

\begin{figure*}%
    \centering
    \includegraphics{fig6.pdf}
    \caption{}
    \label{fig6}
\end{figure*}
\textbf{Fig. 6} Dev non-target responses are reduced due to adaptation in early responders to B.
\textbf{a} The sample network, colored by latency rank, i.e. the rank order of the time to first spike in response to stimulation at site B in any condition. The response develops asymmetrically due to structural properties, not due to interactions with A. Note that non-responding neurons are ranked last and appear dark blue.
\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (dev - msc) and the contrast in $V_{TA}$ at the start of B trials, and associated linear regression. Neurons are colored by latency rank as in panel a. Additionally, the earliest 100 responders are plotted with crosses, while later responders are plotted with dots. Inset histograms show the distributions of the corresponding values across these two categories, and the confidence intervals of their medians, calculated by bootstrap with 10000 iterations.
\textbf{c} Median non-target response contrast (dev - msc) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labelled "late", across networks and stimuli.
\textbf{c} Median contrast (dev - msc) in $V_{TA}$ among the early and late neurons, respectively, across networks and stimuli

@@

Across networks and stimuli, we found even stronger evidence that the early portion of the network drove adaptation to B. We first confirmed that the non-target response and $V_{TA}$ contrasts were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120). In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast (Figure \ref{fig6}c) and the median contrast in $V_{TA}$ at the start of B trials (Figure \ref{fig6}d) within these two portions. We found that the response in the early portion was consistently greater in dev than msc (T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d). The contrast in $V_{TA}$ followed the same pattern, with the early portion exhibiting stronger TA in dev (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).

To sum up, we showed in Figure \ref{fig4} that greater $V_{TA}$ in the portion of the network responding earliest to a stimulus plays an important role in reducing the overall response. Here, we showed that the neurons responding earliest to B exhibit higher $V_{TA}$ in the dev sequence as a consequence of an increased response, in direct contrast to the response and TA levels of the network as a whole.

>>>>

\begin{figure*}%
    \centering
    \includegraphics{fig6.pdf}
    \caption{}
    \label{fig6}
\end{figure*}
\textbf{Fig. 6} Dev non-target responses are reduced due to adaptation in early responders to B.
\textbf{a} The sample network, colored by latency rank, i.e. the rank order of the time to first spike in response to stimulation at site B in any condition. The response develops asymmetrically due to structural properties, not due to interactions with A. Note that non-responding neurons are ranked last and appear dark blue.
\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (dev - msc) and the contrast in $V_{TA}$ at the start of B trials, and associated linear regression. Neurons are colored by latency rank as in panel a. Additionally, the earliest 100 responders are plotted with crosses, while later responders are plotted with dots. Inset histograms show the distributions of the corresponding values across these two categories, and the confidence intervals of their medians, calculated by bootstrap with 10000 iterations.
\textbf{c} Median non-target response contrast (dev - msc) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labelled "late", across networks and stimuli.
\textbf{c} Median contrast (dev - msc) in $V_{TA}$ among the early and late neurons, respectively, across networks and stimuli
\textbf{d} Relationship across networks and stimuli between the median contrast in $V_{TA}$ among the early portion and the median contrast in the response to stimulation in B, and associated linear regression

@@

Across networks and stimuli, we found even stronger evidence that the early portion of the network drove adaptation to B. We first confirmed that the non-target response and $V_{TA}$ contrasts were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120). In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast (Figure \ref{fig6}c) and the median contrast in $V_{TA}$ at the start of B trials (Figure \ref{fig6}d) within these two portions. We found that the response in the early portion was consistently greater in dev than msc (T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d). The contrast in $V_{TA}$ followed the same pattern, with the early portion exhibiting stronger TA in dev (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17). Finally, to show that early $V_{TA}$ was causal in reducing the response to B, we correlated the median contrast in $V_{TA}$ among the early portion of neurons with the network-wide median contrast of the response to stimulation in B, as shown in Figure \ref{fig6}e. Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.

To sum up, we showed in Figure \ref{fig4} that greater $V_{TA}$ in the portion of the network responding earliest to a stimulus plays an important role in reducing the overall response. Here, we showed that the neurons responding earliest to B exhibit higher $V_{TA}$ in the dev sequence as a consequence of an increased response, and that this increase in TA was directly responsible for reducing the response size and TA levels of the network as a whole.
